---
title: "Spotify Data Correlation Matrix"
output: html_notebook
---

# Load the relevant libraries 
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(shiny)
```

# Load the TIDY dataset into a data frame
```{r}
music_data <- read.csv("tidy_spotify_data.csv") 
View(music_data)
```

# List the names of all variables and their classes 
```{r}
sapply(music_data, function(x) class(x))
```

# Select the numerical variables for correlation analysis
numerical_variables <- your_data[, c("NumericVar1", "NumericVar2", "NumericVar3")]
```{r}
num_variables <- music_data[, c("danceability", "mode", "liveness", "popularity", "energy", "speechiness", "valence", "duration_ms", "key", "acousticness", "tempo", "loudness", "instrumentalness", "time_signature")]
```

# Calculate the correlation matrix for numerical variables
```{r}
cor_matrix <- cor(num_variables)
View(cor_matrix)
```

# Set your significance level 
```{r}
alpha <- 0.05
```

# Create a matrix to store p-values
```{r}
p_values_matrix <- matrix(NA, ncol = ncol(cor_matrix), nrow = nrow(cor_matrix))
```

# Loop through the rows and columns of the correlation matrix
```{r}
for (i in 1:ncol(cor_matrix)) {
  for (j in 1:ncol(cor_matrix)) {
    if (i == j) {
      p_values_matrix[i, j] <- 1  # Skip diagonal (correlation with itself)
    } else {
      correlation_test <- cor.test(cor_matrix[i, ], cor_matrix[j, ])
      p_values_matrix[i, j] <- correlation_test$p.value
    }
  }
}
```

# Create a matrix of significant correlations
```{r}
sig_cors <- abs(cor_matrix) * (p_values_matrix < alpha)
View(sig_cors)
```

# Filter out cells equal to 0 (non-significant correlations)
```{r}
# Filter out cells equal to 0 (non-significant correlations)
sig_pairs <- which(sig_cors != 0, arr.ind = TRUE)

# Get the variable names and correlation values involved in each significant correlation
sig_cors_tidy <- data.frame(
  row_var = rownames(sig_cors)[sig_pairs[, 1]],
  col_var = colnames(sig_cors)[sig_pairs[, 2]],
  cor_val = sig_cors[sig_pairs]
)

# Create a unique identifier for each correlation pair
sig_cors_tidy$pair_id <- ifelse(sig_cors_tidy$row_var < sig_cors_tidy$col_var, 
                                paste(sig_cors_tidy$row_var, sig_cors_tidy$col_var, sep = "_"), 
                                paste(sig_cors_tidy$col_var, sig_cors_tidy$row_var, sep = "_"))

# Remove duplicates based on the unique identifier
sig_cors_tidy <- sig_cors_tidy[!duplicated(sig_cors_tidy$pair_id), ]

# Order pairs in descending order (highest correlation value at the top)
sig_cors_tidy <- sig_cors_tidy %>% 
  arrange(desc(cor_val))

# Remove the pair_id column if you don't need it
sig_cors_tidy <- sig_cors_tidy[, -4]

# Print the variable pairs and their correlation values involved in significant correlations
print(sig_cors_tidy)
```


