---
title: "EDA spotify"
author: "Jaishree Ramamoorthi"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

# Load the relevant libraries 
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Load the dataset into a data frame
```{r}
music_data <- read.csv("spotify_music_dataset.csv") 
```

# Data Cleaning
```{r}
music_data <- distinct(music_data)
music_data <- na.omit(music_data)
music_data$explicit <- as.logical(music_data$explicit)
music_data$mode <- as.factor(music_data$mode)
music_data$time_signature <- as.factor(music_data$time_signature)
music_data$key <- as.factor(music_data$key)
```

# Tidy dataset 
```{r}
write.csv(music_data, file = "tidy_spotify_data.csv", row.names = FALSE)
tidy_spotify_data <- read.csv("tidy_spotify_data.csv")
View(tidy_spotify_data)
```

## Exploratory Data Analysis (EDA)

# Summary statistics
```{r}
summary(music_data)
```

# Distribution of popularity
```{r}
ggplot(music_data, aes(x = popularity)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(title = "Distribution of Popularity")
```

# Box plot: Popularity by Genre
```{r}
# Define custom theme with adjusted dimensions and rotated axis labels
custom_theme <- theme(
  plot.title = element_text(size = 16),
  axis.text.x = element_text(angle = 90, hjust = 1, size = 10),
  plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "cm")
)

# Create the box plot with the custom theme
ggplot(music_data, aes(x = track_genre, y = popularity)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Popularity by Genre", x = "Genre", y = "Popularity") +
  coord_flip() +
  scale_y_log10() +
  custom_theme
```
# Scatter plot: Danceability vs. Tempo
```{r}
ggplot(music_data, aes(x = danceability, y = tempo)) +
  geom_point(color = "blue", alpha = 0.7) +
  labs(title = "Danceability vs. Tempo", x = "Danceability", y = "Tempo") +
  theme_minimal()
```
# Scatter plot: Energy vs. Tempo
```{r}
ggplot(music_data, aes(x = energy, y = tempo)) +
  geom_point(color = "green", alpha = 0.7) +
  labs(title = "Energy vs. Tempo", x = "Energy", y = "Tempo") +
  theme_minimal()
```

# Scatter plot: Mode vs. Danceability
```{r}
ggplot(music_data, aes(x = mode, y = danceability)) +
  geom_boxplot(color = "purple", alpha = 0.7) +
  labs(title = "Mode vs. Danceability", x = "Mode", y = "Danceability") +
  theme_minimal()
```

# Scatter plot: Mode vs. Popularity
```{r}
ggplot(music_data, aes(x = mode, y = popularity)) +
  geom_boxplot(color = "red", alpha = 0.7) +
  labs(title = "Mode vs. Popularity", x = "Mode", y = "Popularity") +
  theme_minimal()
```

# Bar Plot: Genre vs. Mode 
```{r}
ggplot(music_data, aes(x = track_genre, fill = mode)) +
  geom_bar(position = "dodge") +
  labs(title = "Genre vs. Mode", x = "Genre", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Box plot: Time Signature vs. Tempo
```{r}
ggplot(music_data, aes(x = time_signature, y = tempo)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Time Signature vs. Tempo", x = "Time Signature", y = "Tempo") +
  theme_minimal()
```


# Box plot: Time Signature vs. Danceability
```{r}
ggplot(music_data, aes(x = time_signature, y = danceability)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Time Signature vs. Danceability", x = "Time Signature", y = "Danceability") +
  theme_minimal()
```

# Scatter plot: Genre vs. Tempo
```{r}
ggplot(music_data, aes(x = track_genre, y = tempo)) +
  geom_point(color = "blue", alpha = 0.7) +
  labs(title = "Genre vs. Tempo", x = "Genre", y = "Tempo") +
  theme_minimal()
```

# Scatter plot: Danceability vs. Speechiness
```{r}
ggplot(music_data, aes(x = danceability, y = speechiness)) +
  geom_point(color = "red", alpha = 0.7) +
  labs(title = "Danceability vs. Speechiness", x = "Danceability", y = "Speechiness") +
  theme_minimal() + coord_flip()
```

# Scatter plot: Energy vs. Loudness
```{r}
ggplot(music_data, aes(x = energy, y = loudness)) +
  geom_point(color = "orange", alpha = 0.7) +
  labs(title = "Energy vs. Loudness", x = "Energy", y = "Loudness") +
  theme_minimal() + coord_flip()
```








